Capabilities, MMIO and Robust Safety

### 2020-03-31 capability-meetings-notes/10-03/main.tex
Some notes on potential clients

\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage{hyperref}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{mathpartir}
\usepackage{tensor}
\usepackage{xspace}
\usepackage[dvipsnames]{xcolor}
\usepackage{iris}
\usepackage{marvosym}
\usepackage{xargs}

@@

\DeclareMathOperator{\initOKo}{init_{OK}}
\DeclareMathOperator{\initOKos}{init^{\X{sem}}_{OK}}
\newcommandx{\initOK}[3][2=r,3=m]{\initOKo(#1,#2,#3)}
\newcommandx{\initOKS}[4][2=r,3=m,4=W]{\initOKos(#1,#2,#3,#4)}
\DeclareMathOperator{\initBCo}{init_{BC}}
\DeclareMathOperator{\initBCos}{init^{\X{sem}}_{BC}}
\newcommandx{\initBC}[2][1=r,2=m]{\initBCo(#1,#2)}
\newcommandx{\initBCS}[3][1=r,2=m,3=W]{\initBCos(#1,#2,#3)}
\DeclareMathOperator{\driverC}{code}

@@

\section{Potential driver clients/properties}
The concrete driver that we described in \texttt{subsec:driverimpl} does not yet enforce any useful
safety properties on the traces that it outputs.
This section describes a couple such properties that we might want to enforce,
modifications to the driver that they require and example clients that make use of our modified driver.

@@

\paragraph{Modelling channel-specific safety properties}

@@

We have to tie this integer $n$ to the number of I/O events
present in the trace $t$, otherwise it will prove impossible to verify our
driver. Concretely, the invariant we will verify our driver under changes from
$\knowInv{\iota}{\exists t\ldotp \tracefrag{t} \ast \pure{|t| \le 1000}}$ to
$\knowInv{\iota}{\exists t\ldotp \tracefrag{t} \ast \pure{|t| \le 1000} \ast
  \exists n\ldotp l_n \mapsto n \ast n = |t| }$. Concretely, this means that
instead of having a pure predicate $P$ in our robust safety theorem, we should
also allow non-pure predicates that entail $P$, since we might have to enforce
other invariants relating to local state over the trace as well.

@@

To make sure we can prove the driver specification, we will have to update the
robust safety theorem slightly to incorporate these changes. It will then look
as follows:

\paragraph{Never performing I/O again after some stop token has been read on $\stdin$}

>>>>

\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage{hyperref}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{mathpartir}
\usepackage{tensor}
\usepackage{xspace}
\usepackage[dvipsnames]{xcolor}
\usepackage{iris}
\usepackage{marvosym}
\usepackage{xargs}
\usepackage{listings}

@@

\DeclareMathOperator{\initOKo}{init_{OK}}
\DeclareMathOperator{\initOKos}{init^{\X{sem}}_{OK}}
\newcommandx{\initOK}[3][2=r,3=m]{\initOKo(#1,#2,#3)}
\newcommandx{\initOKS}[4][2=r,3=m,4=W]{\initOKos(#1,#2,#3,#4)}
\DeclareMathOperator{\initBCo}{init_{BC}}
\DeclareMathOperator{\initBCos}{init^{\X{sem}}_{BC}}
\newcommandx{\initBC}[2][1=r,2=m]{\initBCo(#1,#2)}
\newcommandx{\initBCS}[3][1=r,2=m,3=W]{\initBCos(#1,#2,#3)}
\DeclareMathOperator{\driverC}{code}
\newcommand{\wrap}[1]{\operatorname{wrap}_{\mathrm{#1}}}

@@

\section{Potential driver clients/properties}
The concrete driver that we described in \ref{subsec:driverimpl} does not yet enforce any useful
safety properties on the traces that it outputs.
This section describes a couple such properties that we might want to enforce,
modifications to the driver that they require and example clients that make use of our modified driver.

@@

To only print values $\le 1000$, we need to wrap the driver's read method in a
function that checks this condition, before passing control to the actual read
method.
On an abstract level, given a function $f$, we would have the following wrapper:
\[
  \wrap{\le 1000} f \eqdef \lambda x\ldotp \X{assert}(x \le 1000); f\; x
\]
, and the closure we share with the environment is hence $\wrap{\le
  1000}(\X{read})$ instead of just $\X{read}$.

A trivial misbehaving client could look as follows:
(TODO: ask Dominique for Lau's code on environments)
\begin{lstlisting}
  move r1 1001
  jmp r2
\end{lstlisting}
It simply loads an argument to write and jumps to the write routine.

A well-behaved client might look as follows:
\begin{lstlisting}
  move r1 0
  jmp r2
\end{lstlisting}

\paragraph{Modelling channel-specific safety properties}
Here, we could encode different channels into our integers (in the spirit of
what Frank mentioned), and enforce different safety properties on the different
channels (by ensuring that we can split $P$).
On a conceptual level, this does not require any changes whatsoever.

@@

We have to tie this integer $n$ to the number of I/O events
present in the trace $t$, otherwise it will prove impossible to verify our
driver. Concretely, the invariant we will verify our driver under changes from
$\knowInv{\iota}{\exists t\ldotp \tracefrag{t} \ast \pure{|t| \le 1000}}$ to
$\knowInv{\iota}{\exists t\ldotp \tracefrag{t} \ast \pure{|t| \le 1000} \ast
  \exists n\ldotp l_n \mapsto n \ast n = |t| }$. Concretely, this means that
instead of having a pure predicate $P$ in our robust safety theorem, we should
also allow non-pure predicates $P_{\X{np}}$ that entail $P$, i.e. $\forall t
\ldotp P_{\X{np}}(t) \wand \pure{P(t)} $, since we might have to enforce
other invariants relating to local state over the trace as well.

@@

To make sure we can prove the driver specification, we will have to update the
robust safety theorem slightly to incorporate these changes.

Given a macro instruction to perform loops, a well/mis-behaving client might
look as follows: (TODO: write this out,
and come up with some macro formulation)

\paragraph{Never performing I/O again after some stop token has been read on $\stdin$}
Again, this requires no extra conceptual changes.

\paragraph{Verifying OCP's}
As Arma\"el noted at some point, it should be possible to simulate the
proof-style we see in OCPL by emitting values onto the trace, instead of making
assertions over them. This should give us the same proving power as the OCPL
paper, and allow us to emulate their examples.

\section{Proof sketch}
This section describes the steps involved in proving the robust safety theorem
stated above in some more detail.
